---
title: "Statistical Analysis Plan"
subtitle: "version 1"
date: today
format:
  html:
    css: aside_left_margin.css
    #self-contained: true 
    embed-resources: true
    code-fold: true
    toc: true
    number-sections: true
# execute:
  # cache: true
editor_options:
  chunk_output_type: console
indent: true
indenting: "medium"
editor: 
  markdown: 
    wrap: 72
bibliography: references.bib
csl: american-association-for-cancer-research.csl
---



```{r include=FALSE}
#| message: false
#| warning: false

library(ggplot2)
library(officer)
library(officedown)
library(flextable)
library(data.table)
library(parallel)
library(kableExtra)
library(knitr)
library(lubridate)
library(tictoc)
library(gt)
knitr::opts_chunk$set(echo = FALSE)



ncores<-strtoi(Sys.getenv("SLURM_NTASKS")) #Pick up -ntasks or --n from the environment


```



\


# Hypothesis


Advanced non-small cell lung cancer patients' recommended course of treatments are identified in national cancer .... (NCCN) guideline. The guideline suggests each patient to perform biopsy to identify presence of targetable mutations of certain biomarkers and PDL1 expression to make a informed decision on 1L therapy choice. 

However, the standard guideline may not always be followed due to reasons such as clinician's expert knowledge and ethics. For example, patients with heavy smoking habits mostly develop NSCLC by smoking without any targetable mutation. In this case, chemotherapy alone can be used or if PDL1 expression level is available, any combination of chemotherapy and immunotherapy may be applied 


Combination of chemotherapy or immunotherapy is used for this type of patients in case  chemotherapy is used in absence of PDL1   That is, there usually no targetable mutation, and the PDL1 expression is not useful for treatment. 
In this analysis, we want to estimate the effect of treating advanced non-small cell lung cancer (aNSCLC) patients with 1L therapy without valid PDL1 status or valid targetable mutation status on survival. 

Unfortunately, we do not have clinical trial data where we can derive the causal impact of waiting for 

The standard practice requires either 1 positive mutation status or 2 negative mutation status of biomarkers (ALK, EGFR, BRAF, KRAS, MET, RET, ROS1, NTRK), and percent expression for PDL1 to make informed decision of 1L therapy. In reality, we use certain combination of PDL1 expression and valid gene therapy result but for our simplified version, we will simplify the case so that when a patient observes either a valid targetable mutation status OR valid PDL1 status, then the patient has enough information to inform more precise choice of 1L therapy. Thus, patients who do not have both the PDL1 expression and targetable mutationt status are 

Note: The "valid test" is the one where test result indicating negative or
positive for targetable mutations or PD1L expression. This information
is provided in the dataset in the "BiomarkerStatus" column. If the
column has any one of the following entries, the observation is
considered valid.:

[Check with Wally if the followings are to be used:]{.aside} [- No
interpretation given in report]{.aside} [- Unsuccessful_indeterminate
test]{.aside} [- Unknown]{.aside} [- Variant of Unknown Significance
(VUS) positive]{.aside} [- Results pending]{.aside} [-
Equivocal]{.aside} [- Other result type positive]{.aside} [- Unknown
result type positive]{.aside} [- PD-L1 equivocal]{.aside}

-   Mutation positive

-   PD-L1 positive

-   Rearrangement present

-   Rearrangement positive

-   Amplification positive

-   Protein expression positive

-   PD-L1 positive

-   Mutation negative

-   Negative

-   PD-L1 negative/not detected

-   Rearrangement not present

There are ways to answer this question.

1.  Simple case: fix the maximum time people are allowed to wait to make
    the inference. For example, what is the effect of proceeding
    immediately to 1L within the first 3 weeks since time zero? We would
    estimate risk difference, which is additional survival
    benefit/drawback caused by proceeding immediately to 1L therapy,
    given that patient proceeded immediately to 1L therapy within the
    first 3 weeks. We can change the number of weeks and report causal
    risk difference in a plot in a mock figure @fig-mock_result
2.  general case: do not restrain the number of weeks the 1L therapy
    needs to be initiated by. That is, what is the effect of proceeding
    immediately to 1L? We would estimate risk difference, which is
    additional survival benefit/drawback caused by proceeding
    immediately to 1L therapy.


```{r fig.show="hold", fig.width=6, fig.height=6}
#| label: fig-mock_result
#| code-fold: true
#| echo: false
#| warning: false
#| fig-cap: Mock graph of the result


#generate mock result

this_df<-data.table(
  y=c(0.2,0.1,0.05,0.045,0.040,0.018),
  x=0:5
)
this_df<-data.frame(this_df)
ggplot(this_df, mapping = aes(x=x, y=y))+
  geom_line()+
  geom_point()+
  ylab("Risk difference")+
  xlab("Week (k) from time zero")

```


\

