---
title: "Landmark Analysis - find out patterns of missing"
subtitle: "version 1"
author: "Hyejung Lee <hyejung.lee@utah.edu>"
date: "`r format(Sys.time(), '%a %b %d, %Y %X')`"
always_allow_html: true
output:
  bookdown::github_document2:
    toc: true
    toc_depth: 4
    html_preview: true
  # bookdown::html_document2:
  #   self-contained: true
  #   embed-resources: true
  #   code-fold: true
  #   toc: true
  #   toc_depth: 4
  #   theme: united
  #   toc_float: true
  #   number-sections: true
  # word_document:
  #   toc: true
  #   toc_depth: 3
  #   number_sections: true
# bibliography: references.bib
# csl: nature.csl
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r include=FALSE}
knitr::opts_chunk$set(
  echo=FALSE,
  cache      = TRUE,
  cache.lazy = FALSE,   # turn *off* the lazy‐load DB
  cache.path = "missing_pattern_cache/", # where to put your .rds files
  fig.path   = "missing_pattern_files/"
)

# knitr::knit_hooks$set(chunk = function(x, options) x) #for knitting word_document only

# remotes::install_github('yihui/knitr')
# library(knitr)
library(ggplot2)
library(grid)
library(gridExtra)
library(survival)
library(survminer)
library(officer)
library(officedown)
library(flextable)
library(data.table)
library(parallel)
library(kableExtra)
library(knitr)
library(lubridate)
library(tictoc)
library(gt)
library(gtsummary)
library(tidyverse)
library(DiagrammeR)
library(DiagrammeRsvg)
library(rsvg)






ncores<-strtoi(Sys.getenv("SLURM_NTASKS")) #Pick up -ntasks or --n from the environment
setDTthreads(threads = ncores)

```

```{r pattern-missing-data}
#| echo: false
#| eval: false


#generate data set for identifying patterns of missing
#don't run it in the "knit" because I've already run and 
#saved items as RDS file.

#Read in the data 
ptData_missing<-readRDS("../ptData_time0_diag_v2.rds")
ptData_missing[,advDiag_to_death_weeks:=as.numeric(interval(AdvancedDiagnosisDate,DateOfDeath),"weeks")]
ptData_missing[,advDiag_to_valid_PDL1_week:=as.numeric(interval(AdvancedDiagnosisDate,earliest_valid_ResultDate__on_after_time0_PDL1),"weeks")]
ptData_missing[,advDiag_to_valid_gene_week:=as.numeric(interval(AdvancedDiagnosisDate,earliest_valid_ResultDate__on_after_time0_NoPDL1),"weeks")]
ptData_missing[,advDiag_to_valid_test_week:=pmin(advDiag_to_valid_PDL1_week,advDiag_to_valid_gene_week,na.rm = T)]
ptData_missing[,advDiag_to_1L_week:=as.numeric(interval(AdvancedDiagnosisDate,StartDate),"weeks")]





#Time 0 is 5th week
ptData_missing[,time0:=NULL]
ptData_missing[,time0:=as.IDate(as.Date(AdvancedDiagnosisDate)+weeks(4)+days(1))]


#Now treatment assignment.
ptData_missing[,cohort:="wait"]
ptData_missing[advDiag_to_1L_week<advDiag_to_valid_test_week,cohort:="no_wait"]
ptData_missing[,cohort:=factor(cohort,levels=c("wait","no_wait"), labels=c("wait","no_wait"))]


#generate yearly count of advanced diagnosis date
ptData_missing[,AdvancedDiagnosis_year:=year(AdvancedDiagnosisDate)]
ptData_missing[,AdvancedDiagnosis_year_factor:=factor(AdvancedDiagnosis_year,levels = sort(unique(AdvancedDiagnosis_year), decreasing = F), labels = sort(unique(AdvancedDiagnosis_year), decreasing = F))]

# ptData1[,min(AdvancedDiagnosis_year)] #2011
ptData_missing[,AdvancedDiagnosis_year_rescale:=AdvancedDiagnosis_year-min(AdvancedDiagnosis_year)] #make first one 0.





#Choose the variables that Wally had previously identified for me


lab<-readRDS("../../labs/raw_lab.rds")
vars<-readRDS("../../labs/baseline_covariates.rds")
lab_vars<-unique(vars$variables)

#Exclude carbon dioxide as it changes too frequently.
lab_vars<-lab_vars[lab_vars!="Carbon dioxide"]




lab_values_out<-
lapply(lab_vars,function(this_vars){

  #there can be several different baselines we want to use,
  #choose only one here.
  #set the character string that corresponds to the column name.
  this.baseline<-"AdvancedDiagnosisDate"


  #save the rows we want to use
  this_Test<-vars[variables==this_vars,unique(Test)]

  #Save the subdat
  temp<-lab[Test %in% this_Test,]

  #remove all invalid dataset
  temp<-temp[!is.na(TestResultCleaned) & TestUnitsCleaned!="",]


  #subset the data to what we need.
  these.cols<-c("PatientID",this.baseline)
  temp<-merge.data.table(ptData_missing[,..these.cols],
                         temp[,.(PatientID,TestDate,TestUnitsCleaned,TestResultCleaned)],
                         by="PatientID",
                         all.x=TRUE,
                         allow.cartesian=TRUE)

  setnames(temp, this.baseline,"time0")
  ### Entry Criteria - start ###
  #For baseline value, time windows are:
  #(-12 weeks,4 weeks]
  #(-12 weeks,1 week]

  #generate two
  baseline_dat<-copy(temp)
  baseline_dat[,min_date:=as.IDate(as.Date(time0)- weeks(12))]

  #two different cut off for the baseline
  #Up to 4 weeks after
  #Up to 1 week after
  #Baseline up to 1 week after:
  bdat1<-copy(baseline_dat)
  bdat1[,max_date:=as.IDate(as.Date(time0)+ weeks(1))]

  #Baseline up to 4 weeks after:
  bdat4<-copy(baseline_dat)
  bdat4[,max_date:=as.IDate(as.Date(time0)+ weeks(4))]

  #Save them as a list
  baseline_dat_list<-list(
    "week1"=bdat1,
    "week4"=bdat4
  )

  baseline_dat_list<-
    lapply(baseline_dat_list,function(baseline_dat){
      #Extract only those that are within this time frame
      baseline_dat<-baseline_dat[min_date<=TestDate & max_date>=TestDate,]
      baseline_dat<-baseline_dat[,days_to_time0:=time0-TestDate]
      baseline_dat
    })

baseline_dat_list
})
names(lab_values_out)<-lab_vars

#ECOG
baselineECOG<-readRDS("../../../Data/BaselineECOG.rds")
baselineECOG<-baselineECOG[PatientID %in% ptData_missing$PatientID,]
baselineECOG[ECOGValue=="Unknown",ECOGValue:=NA]
baselineECOG[,ECOGValue:=as.integer(ECOGValue)]
setnames(baselineECOG,"ECOGValue","EcogValue")
setnames(baselineECOG,"ECOGDate","Date")
baselineECOG[,Date:=as.IDate(Date)]


ECOG<-readRDS("../../../Data/ECOG.rds")
ECOG<-ECOG[PatientID %in% ptData_missing$PatientID,]
setnames(ECOG,"EcogDate","Date")
ECOG<-unique(ECOG[,.(PatientID,Date,EcogValue)])
ECOG[,N:=.N,by=.(PatientID,Date)]
ECOG[N>1,] #We will take average
ECOG2<-ECOG[,mean(as.numeric(EcogValue),na.rm=T),by=.(PatientID,Date)]
ECOG2
setnames(ECOG2,"V1","EcogValue")

#get them merged
both_ECOG<-
  merge(unique(baselineECOG[,.(PatientID,Date,EcogValue)]),
        ECOG2[,.(PatientID,Date,EcogValue)],
        by=c("PatientID","Date"),
        all=T)
both_ECOG[,EcogValue.x:=as.numeric(EcogValue.x)]
both_ECOG[is.na(EcogValue.x),EcogValue.x:=EcogValue.y]
both_ECOG[,EcogValue.y:=NULL]
setnames(both_ECOG,"EcogValue.x","EcogValue")
rm(list=c("baselineECOG","ECOG"))
both_ECOG
#Average if more than 1 observed
# both_ECOG<-both_ECOG[,mean(EcogValue, na.rm=T),by=.(PatientID,Date)]

temp<-merge.data.table(ptData_missing[,.(PatientID,AdvancedDiagnosisDate)],
                         both_ECOG,
                         by="PatientID",
                         all.x=TRUE,
                         allow.cartesian=TRUE)


 ### Entry Criteria - start ###
  #For baseline value, time windows are:
  #(-4 months,4 weeks]
  #(-4 months,1 week]

  #generate two
  ECOG_tmp<-copy(temp)
  ECOG_tmp[,min_date:=as.IDate(as.Date(AdvancedDiagnosisDate)%m-% months(4))]

  #two different cut off for the baseline
  #Up to 4 weeks after
  #Up to 1 week after
  #Baseline up to 1 week after:
  bdat1<-copy(ECOG_tmp)
  bdat1[,max_date:=as.IDate(as.Date(AdvancedDiagnosisDate)+ weeks(1))]

  #Baseline up to 4 weeks after:
  bdat4<-copy(ECOG_tmp)
  bdat4[,max_date:=as.IDate(as.Date(AdvancedDiagnosisDate)+ weeks(4))]

  #Save them as a list
  baseline_dat_list<-list(
    "week1"=bdat1,
    "week4"=bdat4
  )

  baseline_dat_list<-
    lapply(baseline_dat_list,function(baseline_dat){
      #Extract only those that are within this time frame
      baseline_dat<-baseline_dat[min_date<=Date & max_date>=Date,]
      baseline_dat<-baseline_dat[,days_to_time0:=AdvancedDiagnosisDate-Date]
      baseline_dat
    })
ECOG<-copy(baseline_dat_list)



#Weight change


#Get baseline weight
vital<-readRDS("../../../Data/Vitals.rds")
vital<-vital[PatientID %in% ptData_missing$PatientID,]

#How many people have weight?
count_weight<-vital[,sum(Test=="body weight"),by=PatientID]
count_weight[,sum(V1==0)]
#224 people have 0 weight. very small number of people. Good.
weight<-vital[Test=="body weight",]

#Want to use kg
#Is there any missing?
weight[is.na(TestResult) & !is.na(TestResultCleaned),.N]
weight[!is.na(TestResult) & is.na(TestResultCleaned),.N]
weight[!is.na(TestResult) & is.na(TestResultCleaned),]
weight[!is.na(TestResult) & is.na(TestResultCleaned),.(TestResult,TestUnits,TestResultCleaned,TestUnitsCleaned)]
#All of TestResult are missing units.
#Just use TestResultCleaned.


#remove all invalid dataset
temp<-weight[!is.na(TestResultCleaned) & TestUnitsCleaned!="",]
temp<-temp[,mean(TestResultCleaned,na.rm=T),by=.(PatientID,TestDate)] #average out if we have more than one observed.
setnames(temp,"V1","TestResultCleaned")
#subset the data to what we need.
temp<-merge.data.table(ptData_missing[,.(PatientID,AdvancedDiagnosisDate)],
                       temp[,.(PatientID,TestDate,TestResultCleaned)],
                       by="PatientID",
                       all.x=TRUE,
                       allow.cartesian=TRUE)

#1 year prior to 1 week or 4 weeks after.
weight_tmp<-copy(temp)
weight_tmp[,min_date:=as.IDate(as.Date(AdvancedDiagnosisDate)%m-% months(12))]

#two different cut off for the baseline
#Up to 4 weeks after
#Up to 1 week after
#Baseline up to 1 week after:
bdat1<-copy(weight_tmp)
bdat1[,max_date:=as.IDate(as.Date(AdvancedDiagnosisDate)+ weeks(1))]

#Baseline up to 4 weeks after:
bdat4<-copy(weight_tmp)
bdat4[,max_date:=as.IDate(as.Date(AdvancedDiagnosisDate)+ weeks(4))]

#Save them as a list
baseline_dat_list<-list(
    "week1"=bdat1,
    "week4"=bdat4
  )

baseline_dat_list<-
    lapply(baseline_dat_list,function(baseline_dat){
      #Extract only those that are within this time frame
      baseline_dat<-baseline_dat[min_date<=TestDate & max_date>=TestDate,]
      baseline_dat<-baseline_dat[,days_to_time0:=AdvancedDiagnosisDate-TestDate]
      baseline_dat
    })
Weight<-copy(baseline_dat_list)
#Average out if more than one is observed.


rm(baseline_dat_list)
rm(weight_tmp)
rm(ECOG_tmp)

saveRDS(Weight, "Weight_pattern_missing.rds")
saveRDS(lab_values_out, "lab_values_out_pattern_missing.rds")
saveRDS(ECOG, "ECOG_pattern_missing.rds")

Weight<-readRDS("Weight_pattern_missing.rds")
lab_values_out<-readRDS("lab_values_out_pattern_missing.rds")
ECOG<-readRDS("ECOG_pattern_missing.rds")

#For each variable, store 12 weeks observation be fore advanced diagnosis date
#and 1 week or 4 weeks after.
all_obs<-ptData_missing[,.(PatientID,AdvancedDiagnosisDate)]
all_obs[,week_12:=as.IDate(as.Date(AdvancedDiagnosisDate)-weeks(12))]
all_obs[,month_4:=as.IDate(as.Date(AdvancedDiagnosisDate)%m-% months(4))]
all_obs[,year_1:=as.IDate(as.Date(AdvancedDiagnosisDate)%m-% months(12))]
all_obs[,week_plus_1:=as.IDate(as.Date(AdvancedDiagnosisDate)+weeks(1))]
all_obs[,week_plus_4:=as.IDate(as.Date(AdvancedDiagnosisDate)+weeks(4))]


#For primary inclusion criteria, we should expand up to 1 week after


#Weights
all_obs_primary_weight<-all_obs[ , .(
  Date = seq(year_1, week_plus_1, by = "days"),
  AdvancedDiagnosisDate     = AdvancedDiagnosisDate          # propagate time0 into each new row
),
by = PatientID
]
all_obs_primary_weight<-
  merge(all_obs_primary_weight,
        Weight$week1[,.(PatientID,TestDate,TestResultCleaned)],
        by.x=c("PatientID","Date"),
        by.y=c("PatientID","TestDate"),
        all.x=T
        )
setnames(all_obs_primary_weight,"TestResultCleaned","weight")


all_obs_secondary_weight<-all_obs[ , .(
  Date = seq(year_1, week_plus_4, by = "days"),
  AdvancedDiagnosisDate     = AdvancedDiagnosisDate          # propagate time0 into each new row
),
by = PatientID
]
all_obs_secondary_weight<-
  merge(all_obs_secondary_weight,
        Weight$week4[,.(PatientID,TestDate,TestResultCleaned)],
        by.x=c("PatientID","Date"),
        by.y=c("PatientID","TestDate"),
        all.x=T
        )
setnames(all_obs_secondary_weight,"TestResultCleaned","weight")

#combine the weights together
weight_missing<-list(
  "week_1" =copy(all_obs_primary_weight),
  "week_4" =copy(all_obs_secondary_weight)
)

rm(list=c("all_obs_primary_weight","all_obs_secondary_weight"))

saveRDS(weight_missing,"weight_missing.rds")


#ECOG
all_obs_primary_ECOG<-all_obs[ , .(
  Date = seq(month_4, week_plus_1, by = "days"),
  AdvancedDiagnosisDate     = AdvancedDiagnosisDate          # propagate time0 into each new row
),
by = PatientID
]
all_obs_primary_ECOG<-
  merge(all_obs_primary_ECOG,
        ECOG$week1[,.(PatientID,Date,EcogValue)],
        by=c("PatientID","Date"),
        all.x=T
  )
setnames(all_obs_primary_ECOG,"EcogValue","ECOG")


all_obs_secondary_ECOG<-all_obs[ , .(
  Date = seq(month_4, week_plus_4, by = "days"),
  AdvancedDiagnosisDate     = AdvancedDiagnosisDate          # propagate time0 into each new row
),
by = PatientID
]
all_obs_secondary_ECOG<-
  merge(all_obs_secondary_ECOG,
        ECOG$week4[,.(PatientID,Date,EcogValue)],
        by=c("PatientID","Date"),
        all.x=T
  )
setnames(all_obs_secondary_ECOG,"EcogValue","ECOG")

#combine the weights together
ECOG_missing<-list(
  "week_1" =copy(all_obs_primary_ECOG),
  "week_4" =copy(all_obs_secondary_ECOG)
)

rm(list=c("all_obs_primary_ECOG","all_obs_secondary_ECOG"))
saveRDS(ECOG_missing,"ECOG_missing.rds")




#Lab values
all_obs_primary<-all_obs[ , .(
  Date = seq(week_12, week_plus_1, by = "days"),
  time0     = AdvancedDiagnosisDate          # propagate time0 into each new row
),
by = PatientID
]

all_obs_secondary<-all_obs[ , .(
  Date = seq(week_12, week_plus_4, by = "days"),
  time0     = AdvancedDiagnosisDate          # propagate time0 into each new row
),
by = PatientID
]


#Merge the values
for(i in 1:length(lab_values_out)){
  print(lab_vars[i])
  #Some lab values have duplicate. Take mean.
  this_lab<-lab_values_out[[i]]$week1[,.(PatientID,TestDate,TestResultCleaned)]
  this_lab<-this_lab[,mean(TestResultCleaned,na.rm=T),by=.(PatientID,TestDate)]
  setnames(this_lab,"V1","TestResultCleaned")

  all_obs_primary<-
  merge(all_obs_primary,
        this_lab,
        by.x=c("PatientID","Date"),
        by.y=c("PatientID","TestDate"),
        all.x=T
        )
  setnames(all_obs_primary,"TestResultCleaned",lab_vars[i])




  this_lab<-lab_values_out[[i]]$week4[,.(PatientID,TestDate,TestResultCleaned)]
  this_lab<-this_lab[,mean(TestResultCleaned,na.rm=T),by=.(PatientID,TestDate)]
  setnames(this_lab,"V1","TestResultCleaned")

  all_obs_secondary<-
  merge(all_obs_secondary,
        this_lab,
        by.x=c("PatientID","Date"),
        by.y=c("PatientID","TestDate"),
        all.x=T
        )
  setnames(all_obs_secondary,"TestResultCleaned",lab_vars[i])
}

labs_missing<-list(
  "week_1" =copy(all_obs_primary),
  "week_4" =copy(all_obs_secondary)
)
saveRDS(labs_missing,"labs_missing.rds")



```

```{r pattern-missing-data-readi}
#| echo: false
#| eval: false



#Read in the data set for exploring missingness
weight_missing<-readRDS("weight_missing.rds")
ECOG_missing<-readRDS("ECOG_missing.rds")
labs_missing<-readRDS("labs_missing.rds")

all_obs_primary<-list(
  "weight"=weight_missing$week_1,
  "ECOG"=ECOG_missing$week_1,
  "lab"=labs_missing$week_1
)

all_obs_secondary<-list(
  "weight"=weight_missing$week_4,
  "ECOG"=ECOG_missing$week_4,
  "lab"=labs_missing$week_4
)

setnames(all_obs_primary$lab,"time0","AdvancedDiagnosisDate")
setnames(all_obs_secondary$lab,"time0","AdvancedDiagnosisDate")





#get variables of interest (VOI)
VOI<-c(
 "Albumin",
 "ECOG",
 "weight"
)

#get auxilliary variables (includes VOI)
auxilliary<-colnames(all_obs_primary$lab)[!colnames(all_obs_primary$lab) %in%c("PatientID","Date","AdvancedDiagnosisDate",VOI)]






#Build dataset for one patient
#For all variable.
#to do so, we have to restrict to -12 weeks because that's the minimum we chose for the variables.
all_obs_primary_week12<-
merge(
  all_obs_primary$lab,
  all_obs_primary$weight[,.(PatientID,Date,weight)],
  by=c("PatientID","Date"),
  all.x=T
)
all_obs_primary_week12<-
merge(
  all_obs_primary_week12,
  all_obs_primary$ECOG[,.(PatientID,Date,ECOG)],
  by=c("PatientID","Date"),
  all.x=T
)
all_obs_primary_week12[,index:=1:.N,by=PatientID]


count_observed<-all_obs_primary_week12[,sum(!is.na(.SD)),by=PatientID,.SDcols=c(VOI,auxilliary)]
count_observed<-count_observed[order(V1,decreasing = T),]  
count_observed[1,] #FC85B4F31A510
#FC85B4F31A510 has the largest number of observation.
#Save this person to work with.
one_obs_primary<-all_obs_primary_week12[PatientID=="FC85B4F31A510",.SD,.SDcols=c(VOI,auxilliary)]



#Secondary inclusion criteria
all_obs_secondary_week12<-
  merge(
    all_obs_secondary$lab,
    all_obs_secondary$weight[,.(PatientID,Date,weight)],
    by=c("PatientID","Date"),
    all.x=T
  )
all_obs_secondary_week12<-
  merge(
    all_obs_secondary_week12,
    all_obs_secondary$ECOG[,.(PatientID,Date,ECOG)],
    by=c("PatientID","Date"),
    all.x=T
  )
all_obs_secondary_week12[,index:=1:.N,by=PatientID]

count_observed<-all_obs_secondary_week12[,sum(!is.na(.SD)),by=PatientID,.SDcols=c(VOI,auxilliary)]
count_observed<-count_observed[order(V1,decreasing = T),]  
count_observed[1,] #F94D0B13A35B4
#F94D0B13A35B4 has the largest number of observation.
#Save this person to work with.
one_obs_secondary<-all_obs_secondary_week12[PatientID=="F94D0B13A35B4",.SD,.SDcols=c(VOI,auxilliary)]



saveRDS(VOI,"VOI.rds")
saveRDS(auxilliary,"auxilliary.rds")

saveRDS(all_obs_primary_week12,"all_obs_primary_week12.rds")
saveRDS(one_obs_primary,"one_obs_primary.rds")

saveRDS(all_obs_secondary_week12,"all_obs_secondary_week12.rds")
saveRDS(one_obs_secondary,"one_obs_secondary.rds")

```

```{r read-data}
#| eval: true
#| echo: false
#| include: false

#Read in the data

VOI<-readRDS("VOI.rds")
auxilliary<-readRDS("auxilliary.rds")

all_obs_primary_week12<-readRDS("all_obs_primary_week12.rds")
one_obs_primary<-readRDS("one_obs_primary.rds")

all_obs_secondary_week12<-readRDS("all_obs_secondary_week12.rds")
one_obs_secondary<-readRDS("one_obs_secondary.rds")


#let's exclude people who are having wrong data 
#in terms of survival time 

#Read in the data 
ptData_missing<-readRDS("../ptData_time0_diag_v2.rds")
ptData_missing[,advDiag_to_death_weeks:=as.numeric(interval(AdvancedDiagnosisDate,DateOfDeath),"weeks")]

# ptData_missing[advDiag_to_death_weeks<=0,table(DeathInd)] 
# #time to death is 0 or less than 0. There are 442 people like that adn they are all censored.
ptData_missing<-ptData_missing[advDiag_to_death_weeks>0,] 


all_obs_primary_week12<-all_obs_primary_week12[PatientID %in% ptData_missing$PatientID,]
all_obs_secondary_week12<-all_obs_secondary_week12[PatientID %in% ptData_missing$PatientID,]


#The below two PatientIDs are what I've chosen for one_obs_primary and one_obs_secondary
"FC85B4F31A510" %in% ptData_missing$PatientID  #TRUE
"F94D0B13A35B4" %in% ptData_missing$PatientID #TRUE

#merge times 
all_obs_primary_week12<-
merge(
  ptData_missing[,.(PatientID,DateOfDeath,DeathInd,earliest_valid_ResultDate__on_after_time0_PDL1,earliest_valid_ResultDate__on_after_time0_NoPDL1,StartDate, SmokingStatus,Gender,RaceEthnicity,age_at_advDiag, Histology)],
  all_obs_primary_week12,
  all=T,
  by="PatientID"
)
all_obs_secondary_week12<-
  merge(
    ptData_missing[,.(PatientID,DateOfDeath,DeathInd,earliest_valid_ResultDate__on_after_time0_PDL1,earliest_valid_ResultDate__on_after_time0_NoPDL1,StartDate, SmokingStatus,Gender,RaceEthnicity,age_at_advDiag, Histology)],
    all_obs_secondary_week12,
    all=T,
    by="PatientID"
  )


one_obs_primary<-
cbind(
  ptData_missing[PatientID=="FC85B4F31A510",.(DateOfDeath,DeathInd,earliest_valid_ResultDate__on_after_time0_PDL1,earliest_valid_ResultDate__on_after_time0_NoPDL1,StartDate, SmokingStatus,Gender,RaceEthnicity,age_at_advDiag, Histology)],
  one_obs_primary
)


one_obs_secondary<-
  cbind(
    ptData_missing[PatientID=="F94D0B13A35B4",.(DateOfDeath,DeathInd,earliest_valid_ResultDate__on_after_time0_PDL1,earliest_valid_ResultDate__on_after_time0_NoPDL1,StartDate, SmokingStatus,Gender,RaceEthnicity,age_at_advDiag, Histology)],
    one_obs_secondary
  )

```

# Context

I've run landmark analysis without imputing covariates (weight change, Albumin, ECOG scores). But by doing so we
lost so many patients! Under the primary inclusion criteria, the sample size decreased from 24690 to 3016, and under the secondary criteria, it decreased from 16397 to 11309. 

Thus, here, we are trying to identify any auxiliary variables that may explain patterns of missingness in the covariates. If we can find any, then we could impute the covariates and re-do the landmark analysis with the imputed data. 

This idea is from the meeting I had with Ben and Tom on 2025/04/30.

\

# Sketch


Let albumin, weight, and ECOG be variables of interest (VOIs). Since the VOIs are longitudinal in nature, we will explore pattenrwe have longitudinal data set, we will look at pattern of missing
in two levels:

1.  Across subjects (missingness over each day), and
2.  Within subjects (missingness by person over time)

For each lab values, the time interval of data collection was chosen
with advanced diagnosis date as the day 0:

-   Albumin (originally from - 6 weeks):
    -   Primary inclusion criteria: $[-12 weeks, 1 week]$
    -   Secondary inclusion criteria: $[-12 weeks, 4 weeks]$
-   ECOG (originally from - 2 months):
    -   Primary inclusion criteria: $[-4 months, 1 week]$
    -   Secondary inclusion criteria: $[-4 months, 4 week]$
-   Weight (originally from - 6 months):
    -   Primary inclusion criteria: $[-1 year, 1 week]$
    -   Secondary inclusion criteria: $[-1 year, 4 weeks]$

\

# Population level missing


For each VOI, we plot proportion of population missing each day. Proportion was calculated with denominator being the number of people alive on the given day.


```{r fig.show="hold", out.height='33%', cache=T}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-population-missing
#| fig-cap: Plots of proportion missing of VOIs for entire population. Proportion was calculated with denominator being the number of people alive on the given number of day since advnaced diagnosis date.




#For each dataset, remove all after censoring date 
labs_missing<-readRDS("labs_missing.rds")



tmp1<-copy(labs_missing)
Albumin_population<-
lapply(tmp1,function(dat){
  
  
  subdat<-
    merge(
      ptData_missing[,.(PatientID,DateOfDeath,DeathInd,earliest_valid_ResultDate__on_after_time0_PDL1,earliest_valid_ResultDate__on_after_time0_NoPDL1,StartDate, SmokingStatus,Gender,RaceEthnicity,age_at_advDiag, Histology)],
      dat,
      all.x=T,
      by="PatientID"
      )
  
  #restrict to before death/censor.
  subdat<-subdat[Date<=DateOfDeath,]
  
  #index for each day
  subdat[,num_day_from_diag:=as.numeric(interval(time0,Date),"days")]
  
  subdat[,mean(is.na(Albumin)),by=num_day_from_diag]

})




tmp2<-data.table(
  group=rep(c("Primary inclusion criteria","Secondary inclusion criteria"), times=c(nrow(Albumin_population[[1]]), nrow(Albumin_population[[2]]))),
  rbind(Albumin_population[[1]],Albumin_population[[2]])
)

ggplot(tmp2, aes(x=num_day_from_diag, y=V1)) +
  geom_line() +
  labs(y="Proportion missing", x="Number of days since advanced diagnosis date") +
  facet_grid(~group)+
  ggtitle("Proportion of missing Albumin")+
  scale_x_continuous(breaks =c(-14*1:6,0,14*1:2) )+
  scale_y_continuous(breaks =seq(0.95,1,by=0.01)) 


# #Zoom in 
# ggplot(tmp2[num_day_from_diag>=-30], aes(x=num_day_from_diag, y=V1)) +
#   geom_line() +
#   labs(y="Proportion missing", x="Number of days since advanced diagnosis date") +
#   facet_grid(~group)+
#   ggtitle("Zoom in to month: Proportion of missing Albumin")+
#   scale_x_continuous(breaks =c(-14*1:6,0,14*1:2) )

#Likewise for ECOG. back to 4 months
ECOG_missing<-readRDS("ECOG_missing.rds")
tmp1<-copy(ECOG_missing)
ECOG_population<-
lapply(tmp1,function(dat){
  
  
  subdat<-
    merge(
      ptData_missing[,.(PatientID,DateOfDeath,DeathInd,earliest_valid_ResultDate__on_after_time0_PDL1,earliest_valid_ResultDate__on_after_time0_NoPDL1,StartDate, SmokingStatus,Gender,RaceEthnicity,age_at_advDiag, Histology)],
      dat,
      all.x=T,
      by="PatientID"
      )
  
  #restrict to before death/censor.
  subdat<-subdat[Date<=DateOfDeath,]
  
  #index for each day
  subdat[,num_day_from_diag:=as.numeric(interval(AdvancedDiagnosisDate,Date),"days")]
  
  subdat[,mean(is.na(ECOG)),by=num_day_from_diag]

})

tmp2<-data.table(
  group=rep(c("Primary inclusion criteria","Secondary inclusion criteria"), times=c(nrow(ECOG_population[[1]]), nrow(ECOG_population[[2]]))),
  rbind(ECOG_population[[1]],ECOG_population[[2]])
)

# tmp2$num_day_from_diag |> range() #-123   28
ggplot(tmp2, aes(x=num_day_from_diag, y=V1)) +
  geom_line() +
  labs(y="Proportion missing", x="Number of days since advanced diagnosis date") +
  facet_grid(~group)+
  ggtitle("Proportion of missing ECOG")+
  # scale_x_continuous(breaks =c(-7*1:16,0,7:1*4) )
  scale_x_continuous(breaks =c(-14*1:8,0,14*1:2) )



#Body weight up to 1 year before
weight_missing<-readRDS("weight_missing.rds")
tmp1<-copy(weight_missing)
weight_population<-
lapply(tmp1,function(dat){
  
  
  subdat<-
    merge(
      ptData_missing[,.(PatientID,DateOfDeath,DeathInd,earliest_valid_ResultDate__on_after_time0_PDL1,earliest_valid_ResultDate__on_after_time0_NoPDL1,StartDate, SmokingStatus,Gender,RaceEthnicity,age_at_advDiag, Histology)],
      dat,
      all.x=T,
      by="PatientID"
      )
  
  #restrict to before death/censor.
  subdat<-subdat[Date<=DateOfDeath,]
  
  #index for each day
  subdat[,num_day_from_diag:=as.numeric(interval(AdvancedDiagnosisDate,Date),"days")]
  
  subdat[,mean(is.na(weight)),by=num_day_from_diag]

})

tmp2<-data.table(
  group=rep(c("Primary inclusion criteria","Secondary inclusion criteria"), times=c(nrow(weight_population[[1]]), nrow(weight_population[[2]]))),
  rbind(weight_population[[1]],weight_population[[2]])
)


ggplot(tmp2, aes(x=num_day_from_diag, y=V1)) +
  geom_line() +
  labs(y="Proportion missing", x="Number of days since advanced diagnosis date") +
  facet_grid(rows=vars(group))+
  ggtitle("Proportion of missing weight")+
  # scale_x_continuous(breaks =c(-7*1:16,0,7:1*4) )
  scale_x_continuous(breaks =c(-14*1:27,0,14*1:2) )

```


Based on the above graphs, it looks like there are patterns of increased number of measurement every 7 days. I wondered if it's the same patient who receives measurement every 7 days. So I looked into the data but that was not the case. 

```{r measure-every-7-days}
#| echo: false
#| eval: false


#Find out If it's the same people who get measurement every 7 days:


Albumin_population$week_1[num_day_from_diag==0,]
Albumin_population$week_1[num_day_from_diag==-3,]
Albumin_population$week_1[num_day_from_diag==-7,]
Albumin_population$week_1[num_day_from_diag==-14,]
Albumin_population$week_1[num_day_from_diag==-18,]


tmp1<-copy(labs_missing)
tmp2<-
lapply(tmp1,function(dat){
  subdat<-
    merge(
      ptData_missing[,.(PatientID,DateOfDeath,DeathInd,earliest_valid_ResultDate__on_after_time0_PDL1,earliest_valid_ResultDate__on_after_time0_NoPDL1,StartDate, SmokingStatus,Gender,RaceEthnicity,age_at_advDiag, Histology)],
      dat,
      all.x=T,
      by="PatientID"
      )
  subdat
  
  #For each person, find out if they have measurement every 7th day 
  #Among the peolpe who have measurement on day 0
  pt<-subdat[!is.na(Albumin) & Date==time0, PatientID]
  
  #Who received every 7 days back for let's day 4 cycles?
  subdat1<-subdat[PatientID %in% pt,]
  subdat1[,time_since_adv_diag_week:=floor(as.numeric(interval(time0,Date),"days")/7)]
  #How many people see observation on every week?
  subdat1<-subdat1[,sum(!is.na(Albumin)>0),by=.(PatientID,time_since_adv_diag_week)]
  subdat1<-subdat1[,mean(V1!=0),by=PatientID]
  prop_weeks_observed_among_observed_on_time0<-subdat1[,table(V1)] #all 1. That means that
  #If you were observed on day 0, then you were observed every week
  
  
  #What about those who don't have record on the day 0?
  pt<-subdat[is.na(Albumin) & Date==time0, PatientID]
  pt<-unique(pt)
  length(pt)#83058
  subdat1<-subdat[PatientID %in% pt,]
  subdat1[,time_since_adv_diag_week:=floor(as.numeric(interval(time0,Date),"days")/7)]
  
  #Make sure everyone has all dates
  check_num_dates<-subdat1[,diff(Date),by=PatientID]
  check_num_dates[,table(V1)] #Only V1. 
  #Good.
  
  #How many people see observation on every week?
  subdat1<-subdat1[,sum(!is.na(Albumin)>0),by=.(PatientID,time_since_adv_diag_week)]
  subdat1<-subdat1[,mean(V1!=0),by=PatientID]
  prop_weeks_observed_among_not_observed_on_time0<-subdat1[,table(V1)] #all 1. That means that
  
  #Return the two proportions
  list(
    "prop_weeks_observed_among_observed_on_time0"=prop_weeks_observed_among_observed_on_time0,
    "prop_weeks_observed_among_not_observed_on_time0"=prop_weeks_observed_among_not_observed_on_time0
  )

})  

tmp2
# $week_1
# $week_1$prop_weeks_observed_among_observed_on_time0
# V1
# 0.0714285714285714  0.142857142857143  0.214285714285714  0.285714285714286  0.357142857142857  0.428571428571429 
#                868                580                235                133                 83                 56 
#                0.5  0.571428571428571  0.642857142857143  0.714285714285714  0.785714285714286  0.857142857142857 
#                 46                 29                 12                 14                  8                  3 
#  0.928571428571429                  1 
#                  3                  2 
# 
# $week_1$prop_weeks_observed_among_not_observed_on_time0
# V1
#                  0 0.0714285714285714  0.142857142857143  0.214285714285714  0.285714285714286  0.357142857142857 
#              59109              16131               4232               1551                738                448 
#  0.428571428571429                0.5  0.571428571428571  0.642857142857143  0.714285714285714  0.785714285714286 
#                340                214                113                 81                 46                 25 
#  0.857142857142857  0.928571428571429 
#                 20                 10 
# 
# 
# $week_4
# $week_4$prop_weeks_observed_among_observed_on_time0
# V1
# 0.0588235294117647  0.117647058823529  0.176470588235294  0.235294117647059  0.294117647058824  0.352941176470588 
#                506                545                372                237                139                 85 
#  0.411764705882353  0.470588235294118  0.529411764705882  0.588235294117647  0.647058823529412  0.705882352941177 
#                 60                 47                 22                 16                 18                 12 
#  0.823529411764706  0.882352941176471  0.941176470588235 
#                  7                  3                  3 
# 
# $week_4$prop_weeks_observed_among_not_observed_on_time0
# V1
#                  0 0.0588235294117647  0.117647058823529  0.176470588235294  0.235294117647059  0.294117647058824 
#              35439              25647              12394               5181               2110                880 
#  0.352941176470588  0.411764705882353  0.470588235294118  0.529411764705882  0.588235294117647  0.647058823529412 
#                494                332                233                129                 86                 57 
#  0.705882352941177  0.764705882352941  0.823529411764706  0.882352941176471  0.941176470588235 
#                 24                 29                 12                  6                  5 


#Based on my calculation, this is what's going in. 
# When I look at people who have observed albumin on time 0, and tried to see if they observed albumin every 7th day, I just did this by finding out whether or not there was any observation weekly. 
#If you look at the output for  $week_4$prop_weeks_observed_among_observed_on_time0, 868 patients are showing proportion = 0.0714285714285714. This means that among 868 patients who observed Albumin on the day of advanced diagnosis, there were only 7.14% of total weeks where the Albumin was observed.
```


\

We present the same graphs but combined the three VOIs in a single single plot. 

```{r fig.show="hold", out.width='49%',  fig.height=6}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-population-missing-comb
#| fig-cap: Plots of proportion missing for entire population 





#We have total 92 days (- 12 weeks , 1 week) for primary
#let's calculate proportion missing for each day of these weeks for 
#each variable.


all_dat<-list(
  "Albumin" = copy(Albumin_population$week_1),
  "ECOG"=copy(ECOG_population$week_1),
  "weight"=copy(weight_population$week_1)
)

all_dat$Albumin[,variable:="Albumin"]
all_dat$ECOG[,variable:="ECOG"]
all_dat$weight[,variable:="weight"]

all_dat<-rbindlist(all_dat)

ggplot(all_dat,
       aes(x=num_day_from_diag, y=V1, color=variable)) +
  geom_line() +
  labs(y="Proportion missing", x="Number of days since advanced diagnosis date") +
  theme_minimal()+
  ggtitle("A) Primary inclusion criteria")




all_dat<-list(
  "Albumin" = copy(Albumin_population$week_4),
  "ECOG"=copy(ECOG_population$week_4),
  "weight"=copy(weight_population$week_4)
)

all_dat$Albumin[,variable:="Albumin"]
all_dat$ECOG[,variable:="ECOG"]
all_dat$weight[,variable:="weight"]

all_dat<-rbindlist(all_dat)

ggplot(all_dat,
       aes(x=num_day_from_diag, y=V1, color=variable)) +
  geom_line() +
  labs(y="Proportion missing", x="Number of days since advanced diagnosis date") +
  theme_minimal()+
  ggtitle("B) Secondary inclusion criteria")

# all_obs_primary_long<-
# melt(all_obs_primary_week12,id.vars = c("PatientID","index"),measure.vars=c(VOI,auxilliary), variable.name = "var", value.name = "val")
# 
# population_summary_primary<-all_obs_primary_long[,mean(is.na(val)),by=.(index,var)]
# setnames(population_summary_primary,"V1","pct_miss")
# 
# ggplot(population_summary_primary[var %in% VOI,],
#        aes(x=index, y=pct_miss, color=var)) +
#   geom_line() +
#   labs(y="Percent missing", x="Day number") +
#   theme_minimal()+
#   ggtitle("A) Primary inclusion criteria")
# 
# 
# 
# 
# #We have total 113 days (- 12 weeks , 4 week) for primary
# #let's calculate proportion missing for each day of these weeks for 
# #each pvariable.
# 
# all_obs_secondary_week12[,index:=1:.N,by=PatientID]
# 
# all_obs_secondary_long<-
#   melt(all_obs_secondary_week12,id.vars = c("PatientID","index"),measure.vars=c(VOI,auxilliary), variable.name = "var", value.name = "val")
# 
# population_summary_secondary<-all_obs_secondary_long[,mean(is.na(val)),by=.(index,var)]
# setnames(population_summary_secondary,"V1","pct_miss")
# 
# 
# ggplot(population_summary_secondary[var %in% VOI,],
#        aes(x=index, y=pct_miss, color=var)) +
#   geom_line() +
#   labs(y="Percent missing", x="Day number") +
#   theme_minimal()+
#   ggtitle("B) Secondary inclusion criteria")

```


\

So... there seems to be some pattern over days. But the proportion of missing is too big. I'm not sure what I can use this information for.



\

I looked at the same graphs for all auxiliary variables, which is shown in Figure \@ref(fig:fig-population-aux). Similar patterns are observed as in VOIs. But again, proportion missing is so large that I don't know if this is worth any.




```{r cache=TRUE,fig.show="hold", out.width='100%',  fig.height=6}
#| echo: false
#| eval: true
#| label: fig-population-aux
#| fig-cap: Proportion of missing observation of all auxiliary variables.



#generate the same graph for all auxiliary variables to see if we can catch the sampe patterns. 


tmp1<-copy(labs_missing)
aux_population<-
lapply(tmp1,function(dat){
  
  
  subdat<-
    merge(
      ptData_missing[,.(PatientID,DateOfDeath,DeathInd,earliest_valid_ResultDate__on_after_time0_PDL1,earliest_valid_ResultDate__on_after_time0_NoPDL1,StartDate, SmokingStatus,Gender,RaceEthnicity,age_at_advDiag, Histology)],
      dat,
      all.x=T,
      by="PatientID"
      )
  
  #restrict to before death/censor.
  subdat<-subdat[Date<=DateOfDeath,]
  
  #index for each day
  subdat[,num_day_from_diag:=as.numeric(interval(time0,Date),"days")]
  
  subdat[,lapply(.SD,function(x)mean(is.na(x))),by=num_day_from_diag, .SDcols = auxilliary]

})

tmp2<-  data.table(
  group=rep(c("Primary inclusion criteria","Secondary inclusion criteria"), times=c(nrow(aux_population[[1]]), nrow(aux_population[[2]]))),
  rbind(aux_population[[1]],aux_population[[2]])
)

tmp3<-tmp2 |> 
melt(id.vars=c("group","num_day_from_diag"),
     measure.vars=auxilliary,
     variable.name="variable",
     value.name = "value")


#Generate label for plot


tmp3[,index:=1:length(value),by=.(group,variable)]
tmp3[,last_obs:=max(index),by=.(group,variable)]
tmp3[,is_last_obs:=(index==last_obs),by=.(group,variable)]

tmp3[,label:=as.character(NA)]
tmp3[is_last_obs==TRUE,label:=variable]



ggplot(tmp3,
       aes(x=num_day_from_diag, y=value, color=variable)) +
  geom_line() +
  facet_grid(cols=vars(group))+
  labs(y="Proportion missing", x="Number of days since advanced diagnosis date") +
  theme_minimal()+
  ggtitle("Proportion of missing auxiliary variables" )
  


```


\

# Subject-level summaries

## Missingness proportion per subject


We will explore proportion of days of missing observation per each patient restricted to 12 weeks before the advanced diagnosis date.

```{r fig.show="hold", out.width='49%',  fig.height=6, fig.width=10}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-subject-level-missing-v1
#| fig-cap: Figure XX. Plots of proportion missing



sub_level_missing_primary1<-all_obs_primary_week12[,lapply(.SD,function(x)mean(is.na(x))),.SDcols = VOI,by=PatientID]

sub_level_missing_secondary1<-all_obs_secondary_week12[,lapply(.SD,function(x)mean(is.na(x))),.SDcols = VOI,by=PatientID]

sub_level_missing_primary1<-
melt(sub_level_missing_primary1,
     id.vars="PatientID",
     measure.vars = VOI,
     variable.name = "voi",
     value.name = "rate")

sub_level_missing_secondary1<-
melt(sub_level_missing_secondary1,
     id.vars="PatientID",
     measure.vars = VOI,
     variable.name = "voi",
     value.name = "rate")


ggplot(sub_level_missing_primary1, mapping = aes(x=rate))+
    geom_histogram(bins=30) +
  # geom_histogram(stat="density") +
  # geom_area(stat = "density", adjust = 0.5)+
  facet_wrap(~voi) +
  labs(x="Proportion of days missing") +
  theme_minimal()+
  ggtitle("A) Primary inclusion criteria")



ggplot(sub_level_missing_secondary1, mapping = aes(x=rate))+
    geom_histogram(bins=30) +
  # geom_histogram(stat="density") +
  # geom_area(stat = "density", adjust = 0.5)+
  facet_wrap(~voi) +
  labs(x="Proportion of days missing") +
  theme_minimal()+
  ggtitle("B) Secondary inclusion criteria")
```

### Correlate subject-level missing rates with auxiliaries

```{r corr-missing-aux-code}
#| echo: false
#| eval: true
#| warning: false
#| message: false



primary_voi_missing_binary<-all_obs_primary_week12[,lapply(.SD,function(x)ifelse(is.na(x),1,0)),.SDcols = VOI]

primary_voi_missing_binary<-cbind(all_obs_primary_week12[,.(PatientID,index)],primary_voi_missing_binary)



secondary_voi_missing_binary<-all_obs_secondary_week12[,lapply(.SD,function(x)ifelse(is.na(x),1,0)),.SDcols = VOI]
secondary_voi_missing_binary<-cbind(all_obs_secondary_week12[,.(PatientID,index)],secondary_voi_missing_binary)




primary_aux_missing_binary<-all_obs_primary_week12[,lapply(.SD,function(x)ifelse(is.na(x),1,0)),.SDcols = auxilliary]
primary_aux_missing_binary<-cbind(all_obs_primary_week12[,.(PatientID,index)],primary_aux_missing_binary)

secondary_aux_missing_binary<-all_obs_secondary_week12[,lapply(.SD,function(x)ifelse(is.na(x),1,0)),.SDcols = auxilliary]
secondary_aux_missing_binary<-cbind(all_obs_secondary_week12[,.(PatientID,index)],secondary_aux_missing_binary)




# join and correlate
primary_missing_binary<-merge(primary_voi_missing_binary,primary_aux_missing_binary,by=c("PatientID","index"))

secondary_missing_binary<-merge(secondary_voi_missing_binary,secondary_aux_missing_binary,by=c("PatientID","index"))



```



```{r results="asis"}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: tbl-corr-missing-aux-primary


library(corrr)


#Primary
cor1<-correlate(primary_missing_binary[,.SD,.SDcols = c(VOI,auxilliary)]) |> focus(Albumin,ECOG,weight) 

data.frame(cor1) |> 
  gt() |> 
  fmt_number(
    decimals=3
  ) |> 
   tab_header(
    title = md("A) Primary inclusion criteria")
    # subtitle = md("`gtcars` is an R dataset")
  ) |>
  tab_caption(caption = md("Table XX. Correleation of variables of interests with all other VOI variables / auxiliary variables.")) |> 
  as_raw_html() |> 
  cat()

```

```{r results='asis'}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: tbl-corr-missing-aux-secondary


#secondary

cor2<-  
correlate(secondary_missing_binary[,.SD,.SDcols = c(VOI,auxilliary)]) |> focus(Albumin,ECOG,weight)

data.frame(cor2) |> 
  gt() |> 
  fmt_number(
    decimals=3
  )|> 
   tab_header(
    title = md("B) Secondary inclusion criteria")
    # subtitle = md("`gtcars` is an R dataset")
  ) |>
  tab_caption(caption = md("Table XX. Correleation of variables of interests with all other VOI variables / auxiliary variables.")) |> 
  as_raw_html() |> 
  cat()



```




```{r fig.show="hold", out.width='49%',  fig.height=6}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-pattern-missing
#| fig-cap: Figure XX. Correlation plot for all lab values (plust ECOG and weight) to identify patterns of missing 



#make a copy that doesn't have date and pateintID


#convert to missing or not missing
all_obs_primary_missing<-copy(all_obs_primary_week12)
all_obs_primary_missing[,c("PatientID","Date","AdvancedDiagnosisDate","index"):=.(NULL)]
all_obs_primary_missing<-all_obs_primary_missing[,lapply(.SD,function(x)ifelse(is.na(x),1,0))]


all_obs_secondary_missing<-copy(all_obs_secondary_week12)
all_obs_secondary_missing[,c("PatientID","Date","AdvancedDiagnosisDate","index"):=.(NULL)]
all_obs_secondary_missing<-all_obs_secondary_missing[,lapply(.SD,function(x)ifelse(is.na(x),1,0))]


#primary exclusion 
DataExplorer::plot_correlation(all_obs_primary_missing,type="continuous", cor_args=list(use="pairwise.complete.obs"), title="A) Primary inclusion criteria") 

#secondary exclusion
DataExplorer::plot_correlation(all_obs_secondary_missing,type="continuous", cor_args=list(use="pairwise.complete.obs"), title="B) Primary inclusion criteria") 
```



\

\

# Subset Calculation


In the entire population, it was hard to identify variables that are associated with each other because there were such high number of missing. Thus, we decided to subset the sample by applying all exclusion criteria we used in the landmark analysis until before excluding patients missing covariates.  those who have test result out and have 1L therapy initiated within either 1 week or 4 weeks after the advanced diagnosis date, for primary and secondary inclusion criteria, respectively. The inclusion criteria are as follows:

-   did not receive valid test result until the day of advanced diagnosis
-   did not initiate 1L therapy before advanced diagnosis date 
-   survived to 4 weeks
-   have received valid test result by 4 weeks


```{r subset, cache=TRUE}
#| eval: true
#| echo: false
#| include: false



ptData_missing<-readRDS("../ptData_time0_diag_v2.rds")

ptData_missing[,advDiag_to_death_weeks:=as.numeric(interval(AdvancedDiagnosisDate,DateOfDeath),"weeks")]
ptData_missing[,advDiag_to_valid_PDL1_week:=as.numeric(interval(AdvancedDiagnosisDate,earliest_valid_ResultDate__on_after_time0_PDL1),"weeks")]
ptData_missing[,advDiag_to_valid_gene_week:=as.numeric(interval(AdvancedDiagnosisDate,earliest_valid_ResultDate__on_after_time0_NoPDL1),"weeks")]
ptData_missing[,advDiag_to_valid_test_week:=pmin(advDiag_to_valid_PDL1_week,advDiag_to_valid_gene_week,na.rm = T)]
ptData_missing[,advDiag_to_1L_week:=as.numeric(interval(AdvancedDiagnosisDate,StartDate),"weeks")]




#Due to data issue, some people are missing censoring time
ptData_missing[DeathInd==0 & advDiag_to_death_weeks<0, .N] #371
pt_exclude<-ptData_missing[DeathInd==0 & advDiag_to_death_weeks<0, PatientID]
#We will exclude them 
ptData_missing<-ptData_missing[!PatientID %in% pt_exclude,]
num_beginning<-nrow(ptData_missing)


#Receive 1L therapy before advanced diagnosis date
ptData_missing[StartDate<AdvancedDiagnosisDate,.N] #994
exclude_pt<-ptData_missing[StartDate<AdvancedDiagnosisDate,PatientID] 
ptData_missing<-ptData_missing[!(PatientID %in% exclude_pt),]




#Receive valid test results before advanced diagnosis date
ptData_missing[observe_valid__before_time0_PDL1==TRUE | observe_valid__before_time0_NoPDL1==TRUE,.N] #6522
exclude_pt<-ptData_missing[observe_valid__before_time0_PDL1==TRUE | observe_valid__before_time0_NoPDL1==TRUE,PatientID] 
ptData_missing<-ptData_missing[!(PatientID %in% exclude_pt),]




#survived up to 4 weeks 
ptData_missing<-ptData_missing[advDiag_to_death_weeks>4,]



#Receive valid test result prior to the end of 4th week
ptData_missing<-ptData_missing[advDiag_to_valid_test_week<=4]

npatients_subset<-nrow(ptData_missing)

```



With the inclusion criteria applied, we have total `r npatients_subset` patients now in our subset for analysis.



As we can see \@ref(fig:fig-population-missing-subset) (which is a same thing as Figure \@ref(fig:fig-population-missing), but in q subset population), there too much missing so that the pattern doesn't become any useful.


```{r fig.show="hold", out.height='33%', cache=T}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-population-missing-subset
#| fig-cap: Plots of proportion missing of VOIs on a subset of population. Proportion was calculated with denominator being the number of people alive on the given number of day since advnaced diagnosis date.



#repeat the same thing on subset. 

#For each dataset, remove all after censoring date 
tmp1<-copy(labs_missing)
Albumin_population<-
lapply(tmp1,function(dat){
  
  
  subdat<-
    merge(
      ptData_missing[,.(PatientID,DateOfDeath)],
      dat,
      all.x=T,
      by="PatientID"
      )
  
  #restrict to before death/censor.
  subdat<-subdat[Date<=DateOfDeath,]
  
  #index for each day
  subdat[,num_day_from_diag:=as.numeric(interval(time0,Date),"days")]
  
  subdat[,mean(is.na(Albumin)),by=num_day_from_diag]

})




tmp2<-data.table(
  group=rep(c("Primary inclusion criteria","Secondary inclusion criteria"), times=c(nrow(Albumin_population[[1]]), nrow(Albumin_population[[2]]))),
  rbind(Albumin_population[[1]],Albumin_population[[2]])
)

ggplot(tmp2, aes(x=num_day_from_diag, y=V1)) +
  geom_line() +
  labs(y="Proportion missing", x="Number of days since advanced diagnosis date") +
  facet_grid(~group)+
  ggtitle("Proportion of missing Albumin")+
  scale_x_continuous(breaks =c(-14*1:6,0,14*1:2) )+
  scale_y_continuous(breaks =seq(0.95,1,by=0.01)) 


# #Zoom in 
# ggplot(tmp2[num_day_from_diag>=-30], aes(x=num_day_from_diag, y=V1)) +
#   geom_line() +
#   labs(y="Proportion missing", x="Number of days since advanced diagnosis date") +
#   facet_grid(~group)+
#   ggtitle("Zoom in to month: Proportion of missing Albumin")+
#   scale_x_continuous(breaks =c(-14*1:6,0,14*1:2) )

#Likewise for ECOG. back to 4 months
tmp1<-copy(ECOG_missing)
ECOG_population<-
lapply(tmp1,function(dat){
  
  
  subdat<-
    merge(
      ptData_missing[,.(PatientID,DateOfDeath)],
      dat,
      all.x=T,
      by="PatientID"
      )
  
  #restrict to before death/censor.
  subdat<-subdat[Date<=DateOfDeath,]
  
  #index for each day
  subdat[,num_day_from_diag:=as.numeric(interval(AdvancedDiagnosisDate,Date),"days")]
  
  subdat[,mean(is.na(ECOG)),by=num_day_from_diag]

})

tmp2<-data.table(
  group=rep(c("Primary inclusion criteria","Secondary inclusion criteria"), times=c(nrow(ECOG_population[[1]]), nrow(ECOG_population[[2]]))),
  rbind(ECOG_population[[1]],ECOG_population[[2]])
)

# tmp2$num_day_from_diag |> range() #-123   28
ggplot(tmp2, aes(x=num_day_from_diag, y=V1)) +
  geom_line() +
  labs(y="Proportion missing", x="Number of days since advanced diagnosis date") +
  facet_grid(~group)+
  ggtitle("Proportion of missing ECOG")+
  # scale_x_continuous(breaks =c(-7*1:16,0,7:1*4) )
  scale_x_continuous(breaks =c(-14*1:8,0,14*1:2) )



#Body weight up to 1 year before
tmp1<-copy(weight_missing)
weight_population<-
lapply(tmp1,function(dat){
  
  
  subdat<-
    merge(
      ptData_missing[,.(PatientID,DateOfDeath)],
      dat,
      all.x=T,
      by="PatientID"
      )
  
  #restrict to before death/censor.
  subdat<-subdat[Date<=DateOfDeath,]
  
  #index for each day
  subdat[,num_day_from_diag:=as.numeric(interval(AdvancedDiagnosisDate,Date),"days")]
  
  subdat[,mean(is.na(weight)),by=num_day_from_diag]

})

tmp2<-data.table(
  group=rep(c("Primary inclusion criteria","Secondary inclusion criteria"), times=c(nrow(weight_population[[1]]), nrow(weight_population[[2]]))),
  rbind(weight_population[[1]],weight_population[[2]])
)


ggplot(tmp2, aes(x=num_day_from_diag, y=V1)) +
  geom_line() +
  labs(y="Proportion missing", x="Number of days since advanced diagnosis date") +
  facet_grid(rows=vars(group))+
  ggtitle("Proportion of missing weight")+
  # scale_x_continuous(breaks =c(-7*1:16,0,7:1*4) )
  scale_x_continuous(breaks =c(-14*1:27,0,14*1:2) )

```


