---
title: "Landmark Analysis - find out pattenrs of missing"
subtitle: "version 1"
author: "Hyejung Lee <hyejung.lee@utah.edu>"
date: "`r format(Sys.time(), '%a %b %d, %Y %X')`"
always_allow_html: true
output:
  bookdown::github_document2:
    toc: true
    toc_depth: 4
    html_preview: true
  # bookdown::html_document2:
  #   self-contained: true
  #   embed-resources: true
  #   code-fold: true
  #   toc: true
  #   toc_depth: 4
  #   theme: united
  #   toc_float: true
  #   number-sections: true
  # word_document:
  #   toc: true
  #   toc_depth: 3
  #   number_sections: true
# bibliography: references.bib
# csl: nature.csl
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r include=FALSE}
knitr::opts_chunk$set(
  echo=FALSE,
  cache      = TRUE,
  cache.lazy = FALSE,   # turn *off* the lazy‐load DB
  cache.path = "missing_pattern_cache/" # where to put your .rds files
)

# knitr::knit_hooks$set(chunk = function(x, options) x) #for knitting word_document only
# knitr::opts_knit$set(root.dir = "./SAP")
# remotes::install_github('yihui/knitr')
# library(knitr)
library(ggplot2)
library(grid)
library(gridExtra)
library(survival)
library(survminer)
library(officer)
library(officedown)
library(flextable)
library(data.table)
library(parallel)
library(kableExtra)
library(knitr)
library(lubridate)
library(tictoc)
library(gt)
library(gtsummary)
library(tidyverse)
library(DiagrammeR)
library(DiagrammeRsvg)
library(rsvg)






ncores<-strtoi(Sys.getenv("SLURM_NTASKS")) #Pick up -ntasks or --n from the environment
setDTthreads(threads = ncores)

```

```{r pattern-missing-data}
#| echo: false
#| eval: false


#generate data set for identifying patterns of missing
#don't run it in the "knit" because I've already run and 
#saved items as RDS file.

#Read in the data 
ptData_missing<-readRDS("../ptData_time0_diag_v2.rds")
ptData_missing[,advDiag_to_death_weeks:=as.numeric(interval(AdvancedDiagnosisDate,DateOfDeath),"weeks")]
ptData_missing[,advDiag_to_valid_PDL1_week:=as.numeric(interval(AdvancedDiagnosisDate,earliest_valid_ResultDate__on_after_time0_PDL1),"weeks")]
ptData_missing[,advDiag_to_valid_gene_week:=as.numeric(interval(AdvancedDiagnosisDate,earliest_valid_ResultDate__on_after_time0_NoPDL1),"weeks")]
ptData_missing[,advDiag_to_valid_test_week:=pmin(advDiag_to_valid_PDL1_week,advDiag_to_valid_gene_week,na.rm = T)]
ptData_missing[,advDiag_to_1L_week:=as.numeric(interval(AdvancedDiagnosisDate,StartDate),"weeks")]





#Time 0 is 5th week
ptData_missing[,time0:=NULL]
ptData_missing[,time0:=as.IDate(as.Date(AdvancedDiagnosisDate)+weeks(4)+days(1))]


#Now treatment assignment.
ptData_missing[,cohort:="wait"]
ptData_missing[advDiag_to_1L_week<advDiag_to_valid_test_week,cohort:="no_wait"]
ptData_missing[,cohort:=factor(cohort,levels=c("wait","no_wait"), labels=c("wait","no_wait"))]


#generate yearly count of advanced diagnosis date
ptData_missing[,AdvancedDiagnosis_year:=year(AdvancedDiagnosisDate)]
ptData_missing[,AdvancedDiagnosis_year_factor:=factor(AdvancedDiagnosis_year,levels = sort(unique(AdvancedDiagnosis_year), decreasing = F), labels = sort(unique(AdvancedDiagnosis_year), decreasing = F))]

# ptData1[,min(AdvancedDiagnosis_year)] #2011
ptData_missing[,AdvancedDiagnosis_year_rescale:=AdvancedDiagnosis_year-min(AdvancedDiagnosis_year)] #make first one 0.





#Choose the variables that Wally had previously identified for me


lab<-readRDS("../../labs/raw_lab.rds")
vars<-readRDS("../../labs/baseline_covariates.rds")
lab_vars<-unique(vars$variables)

#Exclude carbon dioxide as it changes too frequently.
lab_vars<-lab_vars[lab_vars!="Carbon dioxide"]




lab_values_out<-
lapply(lab_vars,function(this_vars){

  #there can be several different baselines we want to use,
  #choose only one here.
  #set the character string that corresponds to the column name.
  this.baseline<-"AdvancedDiagnosisDate"


  #save the rows we want to use
  this_Test<-vars[variables==this_vars,unique(Test)]

  #Save the subdat
  temp<-lab[Test %in% this_Test,]

  #remove all invalid dataset
  temp<-temp[!is.na(TestResultCleaned) & TestUnitsCleaned!="",]


  #subset the data to what we need.
  these.cols<-c("PatientID",this.baseline)
  temp<-merge.data.table(ptData_missing[,..these.cols],
                         temp[,.(PatientID,TestDate,TestUnitsCleaned,TestResultCleaned)],
                         by="PatientID",
                         all.x=TRUE,
                         allow.cartesian=TRUE)

  setnames(temp, this.baseline,"time0")
  ### Entry Criteria - start ###
  #For baseline value, time windows are:
  #(-12 weeks,4 weeks]
  #(-12 weeks,1 week]

  #generate two
  baseline_dat<-copy(temp)
  baseline_dat[,min_date:=as.IDate(as.Date(time0)- weeks(12))]

  #two different cut off for the baseline
  #Up to 4 weeks after
  #Up to 1 week after
  #Baseline up to 1 week after:
  bdat1<-copy(baseline_dat)
  bdat1[,max_date:=as.IDate(as.Date(time0)+ weeks(1))]

  #Baseline up to 4 weeks after:
  bdat4<-copy(baseline_dat)
  bdat4[,max_date:=as.IDate(as.Date(time0)+ weeks(4))]

  #Save them as a list
  baseline_dat_list<-list(
    "week1"=bdat1,
    "week4"=bdat4
  )

  baseline_dat_list<-
    lapply(baseline_dat_list,function(baseline_dat){
      #Extract only those that are within this time frame
      baseline_dat<-baseline_dat[min_date<=TestDate & max_date>=TestDate,]
      baseline_dat<-baseline_dat[,days_to_time0:=time0-TestDate]
      baseline_dat
    })

baseline_dat_list
})
names(lab_values_out)<-lab_vars

#ECOG
baselineECOG<-readRDS("../../../Data/BaselineECOG.rds")
baselineECOG<-baselineECOG[PatientID %in% ptData_missing$PatientID,]
baselineECOG[ECOGValue=="Unknown",ECOGValue:=NA]
baselineECOG[,ECOGValue:=as.integer(ECOGValue)]
setnames(baselineECOG,"ECOGValue","EcogValue")
setnames(baselineECOG,"ECOGDate","Date")
baselineECOG[,Date:=as.IDate(Date)]


ECOG<-readRDS("../../../Data/ECOG.rds")
ECOG<-ECOG[PatientID %in% ptData_missing$PatientID,]
setnames(ECOG,"EcogDate","Date")
ECOG<-unique(ECOG[,.(PatientID,Date,EcogValue)])
ECOG[,N:=.N,by=.(PatientID,Date)]
ECOG[N>1,] #We will take average
ECOG2<-ECOG[,mean(as.numeric(EcogValue),na.rm=T),by=.(PatientID,Date)]
ECOG2
setnames(ECOG2,"V1","EcogValue")

#get them merged
both_ECOG<-
  merge(unique(baselineECOG[,.(PatientID,Date,EcogValue)]),
        ECOG2[,.(PatientID,Date,EcogValue)],
        by=c("PatientID","Date"),
        all=T)
both_ECOG[,EcogValue.x:=as.numeric(EcogValue.x)]
both_ECOG[is.na(EcogValue.x),EcogValue.x:=EcogValue.y]
both_ECOG[,EcogValue.y:=NULL]
setnames(both_ECOG,"EcogValue.x","EcogValue")
rm(list=c("baselineECOG","ECOG"))
both_ECOG
#Average if more than 1 observed
# both_ECOG<-both_ECOG[,mean(EcogValue, na.rm=T),by=.(PatientID,Date)]

temp<-merge.data.table(ptData_missing[,.(PatientID,AdvancedDiagnosisDate)],
                         both_ECOG,
                         by="PatientID",
                         all.x=TRUE,
                         allow.cartesian=TRUE)


 ### Entry Criteria - start ###
  #For baseline value, time windows are:
  #(-4 months,4 weeks]
  #(-4 months,1 week]

  #generate two
  ECOG_tmp<-copy(temp)
  ECOG_tmp[,min_date:=as.IDate(as.Date(AdvancedDiagnosisDate)%m-% months(4))]

  #two different cut off for the baseline
  #Up to 4 weeks after
  #Up to 1 week after
  #Baseline up to 1 week after:
  bdat1<-copy(ECOG_tmp)
  bdat1[,max_date:=as.IDate(as.Date(AdvancedDiagnosisDate)+ weeks(1))]

  #Baseline up to 4 weeks after:
  bdat4<-copy(ECOG_tmp)
  bdat4[,max_date:=as.IDate(as.Date(AdvancedDiagnosisDate)+ weeks(4))]

  #Save them as a list
  baseline_dat_list<-list(
    "week1"=bdat1,
    "week4"=bdat4
  )

  baseline_dat_list<-
    lapply(baseline_dat_list,function(baseline_dat){
      #Extract only those that are within this time frame
      baseline_dat<-baseline_dat[min_date<=Date & max_date>=Date,]
      baseline_dat<-baseline_dat[,days_to_time0:=AdvancedDiagnosisDate-Date]
      baseline_dat
    })
ECOG<-copy(baseline_dat_list)



#Weight change


#Get baseline weight
vital<-readRDS("../../../Data/Vitals.rds")
vital<-vital[PatientID %in% ptData_missing$PatientID,]

#How many people have weight?
count_weight<-vital[,sum(Test=="body weight"),by=PatientID]
count_weight[,sum(V1==0)]
#224 people have 0 weight. very small number of people. Good.
weight<-vital[Test=="body weight",]

#Want to use kg
#Is there any missing?
weight[is.na(TestResult) & !is.na(TestResultCleaned),.N]
weight[!is.na(TestResult) & is.na(TestResultCleaned),.N]
weight[!is.na(TestResult) & is.na(TestResultCleaned),]
weight[!is.na(TestResult) & is.na(TestResultCleaned),.(TestResult,TestUnits,TestResultCleaned,TestUnitsCleaned)]
#All of TestResult are missing units.
#Just use TestResultCleaned.


#remove all invalid dataset
temp<-weight[!is.na(TestResultCleaned) & TestUnitsCleaned!="",]
temp<-temp[,mean(TestResultCleaned,na.rm=T),by=.(PatientID,TestDate)] #average out if we have more than one observed.
setnames(temp,"V1","TestResultCleaned")
#subset the data to what we need.
temp<-merge.data.table(ptData_missing[,.(PatientID,AdvancedDiagnosisDate)],
                       temp[,.(PatientID,TestDate,TestResultCleaned)],
                       by="PatientID",
                       all.x=TRUE,
                       allow.cartesian=TRUE)

#1 year prior to 1 week or 4 weeks after.
weight_tmp<-copy(temp)
weight_tmp[,min_date:=as.IDate(as.Date(AdvancedDiagnosisDate)%m-% months(12))]

#two different cut off for the baseline
#Up to 4 weeks after
#Up to 1 week after
#Baseline up to 1 week after:
bdat1<-copy(weight_tmp)
bdat1[,max_date:=as.IDate(as.Date(AdvancedDiagnosisDate)+ weeks(1))]

#Baseline up to 4 weeks after:
bdat4<-copy(weight_tmp)
bdat4[,max_date:=as.IDate(as.Date(AdvancedDiagnosisDate)+ weeks(4))]

#Save them as a list
baseline_dat_list<-list(
    "week1"=bdat1,
    "week4"=bdat4
  )

baseline_dat_list<-
    lapply(baseline_dat_list,function(baseline_dat){
      #Extract only those that are within this time frame
      baseline_dat<-baseline_dat[min_date<=TestDate & max_date>=TestDate,]
      baseline_dat<-baseline_dat[,days_to_time0:=AdvancedDiagnosisDate-TestDate]
      baseline_dat
    })
Weight<-copy(baseline_dat_list)
#Average out if more than one is observed.


rm(baseline_dat_list)
rm(weight_tmp)
rm(ECOG_tmp)

saveRDS(Weight, "Weight_pattern_missing.rds")
saveRDS(lab_values_out, "lab_values_out_pattern_missing.rds")
saveRDS(ECOG, "ECOG_pattern_missing.rds")

Weight<-readRDS("Weight_pattern_missing.rds")
lab_values_out<-readRDS("lab_values_out_pattern_missing.rds")
ECOG<-readRDS("ECOG_pattern_missing.rds")

#For each variable, store 12 weeks observation be fore advanced diagnosis date
#and 1 week or 4 weeks after.
all_obs<-ptData_missing[,.(PatientID,AdvancedDiagnosisDate)]
all_obs[,week_12:=as.IDate(as.Date(AdvancedDiagnosisDate)-weeks(12))]
all_obs[,month_4:=as.IDate(as.Date(AdvancedDiagnosisDate)%m-% months(4))]
all_obs[,year_1:=as.IDate(as.Date(AdvancedDiagnosisDate)%m-% months(12))]
all_obs[,week_plus_1:=as.IDate(as.Date(AdvancedDiagnosisDate)+weeks(1))]
all_obs[,week_plus_4:=as.IDate(as.Date(AdvancedDiagnosisDate)+weeks(4))]


#For primary inclusion criteria, we should expand up to 1 week after


#Weights
all_obs_primary_weight<-all_obs[ , .(
  Date = seq(year_1, week_plus_1, by = "days"),
  AdvancedDiagnosisDate     = AdvancedDiagnosisDate          # propagate time0 into each new row
),
by = PatientID
]
all_obs_primary_weight<-
  merge(all_obs_primary_weight,
        Weight$week1[,.(PatientID,TestDate,TestResultCleaned)],
        by.x=c("PatientID","Date"),
        by.y=c("PatientID","TestDate"),
        all.x=T
        )
setnames(all_obs_primary_weight,"TestResultCleaned","weight")


all_obs_secondary_weight<-all_obs[ , .(
  Date = seq(year_1, week_plus_4, by = "days"),
  AdvancedDiagnosisDate     = AdvancedDiagnosisDate          # propagate time0 into each new row
),
by = PatientID
]
all_obs_secondary_weight<-
  merge(all_obs_secondary_weight,
        Weight$week4[,.(PatientID,TestDate,TestResultCleaned)],
        by.x=c("PatientID","Date"),
        by.y=c("PatientID","TestDate"),
        all.x=T
        )
setnames(all_obs_secondary_weight,"TestResultCleaned","weight")

#combine the weights together
weight_missing<-list(
  "week_1" =copy(all_obs_primary_weight),
  "week_4" =copy(all_obs_secondary_weight)
)

rm(list=c("all_obs_primary_weight","all_obs_secondary_weight"))

saveRDS(weight_missing,"weight_missing.rds")


#ECOG
all_obs_primary_ECOG<-all_obs[ , .(
  Date = seq(month_4, week_plus_1, by = "days"),
  AdvancedDiagnosisDate     = AdvancedDiagnosisDate          # propagate time0 into each new row
),
by = PatientID
]
all_obs_primary_ECOG<-
  merge(all_obs_primary_ECOG,
        ECOG$week1[,.(PatientID,Date,EcogValue)],
        by=c("PatientID","Date"),
        all.x=T
  )
setnames(all_obs_primary_ECOG,"EcogValue","ECOG")


all_obs_secondary_ECOG<-all_obs[ , .(
  Date = seq(month_4, week_plus_4, by = "days"),
  AdvancedDiagnosisDate     = AdvancedDiagnosisDate          # propagate time0 into each new row
),
by = PatientID
]
all_obs_secondary_ECOG<-
  merge(all_obs_secondary_ECOG,
        ECOG$week4[,.(PatientID,Date,EcogValue)],
        by=c("PatientID","Date"),
        all.x=T
  )
setnames(all_obs_secondary_ECOG,"EcogValue","ECOG")

#combine the weights together
ECOG_missing<-list(
  "week_1" =copy(all_obs_primary_ECOG),
  "week_4" =copy(all_obs_secondary_ECOG)
)

rm(list=c("all_obs_primary_ECOG","all_obs_secondary_ECOG"))
saveRDS(ECOG_missing,"ECOG_missing.rds")




#Lab values
all_obs_primary<-all_obs[ , .(
  Date = seq(week_12, week_plus_1, by = "days"),
  time0     = AdvancedDiagnosisDate          # propagate time0 into each new row
),
by = PatientID
]

all_obs_secondary<-all_obs[ , .(
  Date = seq(week_12, week_plus_4, by = "days"),
  time0     = AdvancedDiagnosisDate          # propagate time0 into each new row
),
by = PatientID
]


#Merge the values
for(i in 1:length(lab_values_out)){
  print(lab_vars[i])
  #Some lab values have duplicate. Take mean.
  this_lab<-lab_values_out[[i]]$week1[,.(PatientID,TestDate,TestResultCleaned)]
  this_lab<-this_lab[,mean(TestResultCleaned,na.rm=T),by=.(PatientID,TestDate)]
  setnames(this_lab,"V1","TestResultCleaned")

  all_obs_primary<-
  merge(all_obs_primary,
        this_lab,
        by.x=c("PatientID","Date"),
        by.y=c("PatientID","TestDate"),
        all.x=T
        )
  setnames(all_obs_primary,"TestResultCleaned",lab_vars[i])




  this_lab<-lab_values_out[[i]]$week4[,.(PatientID,TestDate,TestResultCleaned)]
  this_lab<-this_lab[,mean(TestResultCleaned,na.rm=T),by=.(PatientID,TestDate)]
  setnames(this_lab,"V1","TestResultCleaned")

  all_obs_secondary<-
  merge(all_obs_secondary,
        this_lab,
        by.x=c("PatientID","Date"),
        by.y=c("PatientID","TestDate"),
        all.x=T
        )
  setnames(all_obs_secondary,"TestResultCleaned",lab_vars[i])
}

labs_missing<-list(
  "week_1" =copy(all_obs_primary),
  "week_4" =copy(all_obs_secondary)
)
saveRDS(labs_missing,"labs_missing.rds")



```

```{r pattern-missing-data-readi}
#| echo: false
#| eval: false



#Read in the data set for exploring missingness
weight_missing<-readRDS("weight_missing.rds")
ECOG_missing<-readRDS("ECOG_missing.rds")
labs_missing<-readRDS("labs_missing.rds")

all_obs_primary<-list(
  "weight"=weight_missing$week_1,
  "ECOG"=ECOG_missing$week_1,
  "lab"=labs_missing$week_1
)

all_obs_secondary<-list(
  "weight"=weight_missing$week_4,
  "ECOG"=ECOG_missing$week_4,
  "lab"=labs_missing$week_4
)

setnames(all_obs_primary$lab,"time0","AdvancedDiagnosisDate")
setnames(all_obs_secondary$lab,"time0","AdvancedDiagnosisDate")





#get variables of interest (VOI)
VOI<-c(
 "Albumin",
 "ECOG",
 "weight"
)

#get auxilliary variables (includes VOI)
auxilliary<-colnames(all_obs_primary$lab)[!colnames(all_obs_primary$lab) %in%c("PatientID","Date","AdvancedDiagnosisDate",VOI)]






#Build dataset for one patient
#For all variable.
#to do so, we have to restrict to -12 weeks because that's the minimum we chose for the variables.
all_obs_primary_week12<-
merge(
  all_obs_primary$lab,
  all_obs_primary$weight[,.(PatientID,Date,weight)],
  by=c("PatientID","Date"),
  all.x=T
)
all_obs_primary_week12<-
merge(
  all_obs_primary_week12,
  all_obs_primary$ECOG[,.(PatientID,Date,ECOG)],
  by=c("PatientID","Date"),
  all.x=T
)
all_obs_primary_week12[,index:=1:.N,by=PatientID]


count_observed<-all_obs_primary_week12[,sum(!is.na(.SD)),by=PatientID,.SDcols=c(VOI,auxilliary)]
count_observed<-count_observed[order(V1,decreasing = T),]  
count_observed[1,] #FC85B4F31A510
#FC85B4F31A510 has the largest number of observation.
#Save this person to work with.
one_obs_primary<-all_obs_primary_week12[PatientID=="FC85B4F31A510",.SD,.SDcols=c(VOI,auxilliary)]



#Secondary inclusion criteria
all_obs_secondary_week12<-
  merge(
    all_obs_secondary$lab,
    all_obs_secondary$weight[,.(PatientID,Date,weight)],
    by=c("PatientID","Date"),
    all.x=T
  )
all_obs_secondary_week12<-
  merge(
    all_obs_secondary_week12,
    all_obs_secondary$ECOG[,.(PatientID,Date,ECOG)],
    by=c("PatientID","Date"),
    all.x=T
  )
all_obs_secondary_week12[,index:=1:.N,by=PatientID]

count_observed<-all_obs_secondary_week12[,sum(!is.na(.SD)),by=PatientID,.SDcols=c(VOI,auxilliary)]
count_observed<-count_observed[order(V1,decreasing = T),]  
count_observed[1,] #F94D0B13A35B4
#F94D0B13A35B4 has the largest number of observation.
#Save this person to work with.
one_obs_secondary<-all_obs_secondary_week12[PatientID=="F94D0B13A35B4",.SD,.SDcols=c(VOI,auxilliary)]



saveRDS(VOI,"VOI.rds")
saveRDS(auxilliary,"auxilliary.rds")

saveRDS(all_obs_primary_week12,"all_obs_primary_week12.rds")
saveRDS(one_obs_primary,"one_obs_primary.rds")

saveRDS(all_obs_secondary_week12,"all_obs_secondary_week12.rds")
saveRDS(one_obs_secondary,"one_obs_secondary.rds")

```

```{r read-data}
#| eval: true
#| echo: false

#Read in the data

VOI<-readRDS("VOI.rds")
auxilliary<-readRDS("auxilliary.rds")

all_obs_primary_week12<-readRDS("all_obs_primary_week12.rds")
one_obs_primary<-readRDS("one_obs_primary.rds")

all_obs_secondary_week12<-readRDS("all_obs_secondary_week12.rds")
one_obs_secondary<-readRDS("one_obs_secondary.rds")


```

# Context

I've run landmark analysis without imputing any data. But by doing so we
lost so much samples. Thus, we should look into patterns of missing to
check if we can impute or not. This idea is from the meeting I had with
Ben and Tom on 2025/04/30

# Sketch

Since we have longitudinal data set, we will look at pattern of missing
in two levels:

1.  Across subjects (missingness over each visit), and
2.  Within subjects (missingness by person over time)

For each lab values, the time interval of data collection was chosen
with advanced diagnosis date as the day 0:

-   Albumin (originally from - 6 weeks):
    -   Primary inclusion criteria: $[-12 weeks, 1 week]$
    -   Secondary inclusion criteria: $[-12 weeks, 4 weeks]$
-   ECOG (originally from - 2 months):
    -   Primary inclusion criteria: $[-4 months, 1 week]$
    -   Secondary inclusion criteria: $[-4 months, 4 week]$
-   Weight (originally from - 6 months):
    -   Primary inclusion criteria: $[-1 year, 1 week]$
    -   Secondary inclusion criteria: $[-1 year, 4 weeks]$

\

# Population level missing

```{r fig.show="hold", out.width='49%',  fig.height=6}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-population-missing
#| fig-cap: Plots of proportion missing for entire population 





#We have total 92 days (- 12 weeks , 1 week) for primary
#let's calculate proportion missing for each day of these weeks for 
#each variable.



all_obs_primary_long<-
melt(all_obs_primary_week12,id.vars = c("PatientID","index"),measure.vars=c(VOI,auxilliary), variable.name = "var", value.name = "val")

population_summary_primary<-all_obs_primary_long[,mean(is.na(val)),by=.(index,var)]
setnames(population_summary_primary,"V1","pct_miss")

ggplot(population_summary_primary[var %in% VOI,],
       aes(x=index, y=pct_miss, color=var)) +
  geom_line() +
  labs(y="Percent missing", x="Day number") +
  theme_minimal()+
  ggtitle("A) Primary inclusion criteria")




#We have total 113 days (- 12 weeks , 4 week) for primary
#let's calculate proportion missing for each day of these weeks for 
#each pvariable.

all_obs_secondary_week12[,index:=1:.N,by=PatientID]

all_obs_secondary_long<-
  melt(all_obs_secondary_week12,id.vars = c("PatientID","index"),measure.vars=c(VOI,auxilliary), variable.name = "var", value.name = "val")

population_summary_secondary<-all_obs_secondary_long[,mean(is.na(val)),by=.(index,var)]
setnames(population_summary_secondary,"V1","pct_miss")


ggplot(population_summary_secondary[var %in% VOI,],
       aes(x=index, y=pct_miss, color=var)) +
  geom_line() +
  labs(y="Percent missing", x="Day number") +
  theme_minimal()+
  ggtitle("B) Secondary inclusion criteria")

```

\

# Subject-level summaries

## Missingness proportion per subject

Again, we will restrict to week 12 before advanced diagnosis date even
though we have up to 4 months and 1 year before the advanced diagnosis
date for ECOG and weights.

```{r fig.show="hold", out.width='49%',  fig.height=6, fig.width=10}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-subject-level-missing-v1
#| fig-cap: Figure XX. Plots of proportion missing



sub_level_missing_primary1<-all_obs_primary_week12[,lapply(.SD,function(x)mean(is.na(x))),.SDcols = VOI,by=PatientID]

sub_level_missing_secondary1<-all_obs_secondary_week12[,lapply(.SD,function(x)mean(is.na(x))),.SDcols = VOI,by=PatientID]

sub_level_missing_primary1<-
melt(sub_level_missing_primary1,
     id.vars="PatientID",
     measure.vars = VOI,
     variable.name = "voi",
     value.name = "rate")

sub_level_missing_secondary1<-
melt(sub_level_missing_secondary1,
     id.vars="PatientID",
     measure.vars = VOI,
     variable.name = "voi",
     value.name = "rate")


ggplot(sub_level_missing_primary1, mapping = aes(x=rate))+
    geom_histogram(bins=30) +
  # geom_histogram(stat="density") +
  # geom_area(stat = "density", adjust = 0.5)+
  facet_wrap(~voi) +
  labs(x="Proportion of days missing") +
  theme_minimal()+
  ggtitle("A) Primary inclusion criteria")



ggplot(sub_level_missing_secondary1, mapping = aes(x=rate))+
    geom_histogram(bins=30) +
  # geom_histogram(stat="density") +
  # geom_area(stat = "density", adjust = 0.5)+
  facet_wrap(~voi) +
  labs(x="Proportion of days missing") +
  theme_minimal()+
  ggtitle("B) Secondary inclusion criteria")
```

### Correlate subject-level missing rates with auxiliaries

```{r corr-missing-aux-code}
#| echo: false
#| eval: true
#| warning: false
#| message: false



primary_voi_missing_binary<-all_obs_primary_week12[,lapply(.SD,function(x)ifelse(is.na(x),1,0)),.SDcols = VOI]

primary_voi_missing_binary<-cbind(all_obs_primary_week12[,.(PatientID,index)],primary_voi_missing_binary)



secondary_voi_missing_binary<-all_obs_secondary_week12[,lapply(.SD,function(x)ifelse(is.na(x),1,0)),.SDcols = VOI]
secondary_voi_missing_binary<-cbind(all_obs_secondary_week12[,.(PatientID,index)],secondary_voi_missing_binary)




primary_aux_missing_binary<-all_obs_primary_week12[,lapply(.SD,function(x)ifelse(is.na(x),1,0)),.SDcols = auxilliary]
primary_aux_missing_binary<-cbind(all_obs_primary_week12[,.(PatientID,index)],primary_aux_missing_binary)

secondary_aux_missing_binary<-all_obs_secondary_week12[,lapply(.SD,function(x)ifelse(is.na(x),1,0)),.SDcols = auxilliary]
secondary_aux_missing_binary<-cbind(all_obs_secondary_week12[,.(PatientID,index)],secondary_aux_missing_binary)




# join and correlate
primary_missing_binary<-merge(primary_voi_missing_binary,primary_aux_missing_binary,by=c("PatientID","index"))

secondary_missing_binary<-merge(secondary_voi_missing_binary,secondary_aux_missing_binary,by=c("PatientID","index"))



```



```{r results="asis"}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: tbl-corr-missing-aux-primary


library(corrr)


#Primary
cor1<-correlate(primary_missing_binary[,.SD,.SDcols = c(VOI,auxilliary)]) |> focus(Albumin,ECOG,weight) 

data.frame(cor1) |> 
  gt() |> 
  fmt_number(
    decimals=3
  ) |> 
   tab_header(
    title = md("A) Primary inclusion criteria")
    # subtitle = md("`gtcars` is an R dataset")
  ) |>
  tab_caption(caption = md("Table XX. Correleation of variables of interests with all other VOI variables / auxiliary variables.")) |> 
  as_raw_html() |> 
  cat()

```

```{r results='asis'}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: tbl-corr-missing-aux-secondary


#secondary

cor2<-  
correlate(secondary_missing_binary[,.SD,.SDcols = c(VOI,auxilliary)]) |> focus(Albumin,ECOG,weight)

data.frame(cor2) |> 
  gt() |> 
  fmt_number(
    decimals=3
  )|> 
   tab_header(
    title = md("B) Secondary inclusion criteria")
    # subtitle = md("`gtcars` is an R dataset")
  ) |>
  tab_caption(caption = md("Table XX. Correleation of variables of interests with all other VOI variables / auxiliary variables.")) |> 
  as_raw_html() |> 
  cat()



```




```{r fig.show="hold", out.width='49%',  fig.height=6}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-pattern-missing
#| fig-cap: Figure XX. Correlation plot for all lab values (plust ECOG and weight) to identify patterns of missing 



#make a copy that doesn't have date and pateintID


#convert to missing or not missing
all_obs_primary_missing<-copy(all_obs_primary_week12)
all_obs_primary_missing[,c("PatientID","Date","AdvancedDiagnosisDate","index"):=.(NULL)]
all_obs_primary_missing<-all_obs_primary_missing[,lapply(.SD,function(x)ifelse(is.na(x),1,0))]


all_obs_secondary_missing<-copy(all_obs_secondary_week12)
all_obs_secondary_missing[,c("PatientID","Date","AdvancedDiagnosisDate","index"):=.(NULL)]
all_obs_secondary_missing<-all_obs_secondary_missing[,lapply(.SD,function(x)ifelse(is.na(x),1,0))]


#primary exclusion 
DataExplorer::plot_correlation(all_obs_primary_missing,type="continuous", cor_args=list(use="pairwise.complete.obs"), title="A) Primary inclusion criteria") 

#secondary exclusion
DataExplorer::plot_correlation(all_obs_secondary_missing,type="continuous", cor_args=list(use="pairwise.complete.obs"), title="B) Primary inclusion criteria") 
```
